---
title: "Intro to importing data"
author: "Sungbae Chun"
date: "5/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 5 types of data

 - flat files
 - data from excel
 - databases
 - web
 - statistical software (SAS, STATA, SPSS)


## Importing data from flat files

### Using utils

Importing to dataframe from csv files: utils - read.csv
<br>
`read.csv("<path>", args...)`<br>
`path = file.path("~","data","my_data.csv")`

set `header=F` if data starts withough labels

```{r}
path = file.path("~","Courses","ds_r","im_cl_data","gpa.csv")
gpa2020s = read.csv(path,stringsAsFactors = F)
gpa2020s
str(gpa2020s)
```
<br><br>
**Dealing with delimited files**<br>
`read.delim("<path",sep="<your_delim>")`
<br>
`df[which.min(df$col),]` selects min<br><br>
**Explicit class**<br>
`read.delim(<path>, colClasses = <str_vector of classnames>)`
<br>Columns with `NULL` colclass will be skipped

<br>


### Using readr

Why use readr instead of utils?
 - Fast, easy to use, consistent
 - utils: complex expr, slow
```{r}
library(readr)
```

```{r}
gpa = read_csv("./gpa.csv",col_types = "cccdd")
#col_types, c - char, d - double, i - integer, l - logical, _ - skip, ? - guess
# t - time, T - date time, D - date, n - number
#col_names = <str_vec> to specify col names
gpa$grade = factor(x=gpa$grade, ordered = T, levels = c("B","B+","A-","A"))

ch = col_character()
f = col_factor(ordered = T, levels = c("C+","B-","B","B+","A-","A"))
i = col_integer()
d = col_double()
gpa = read_csv('./gpa.csv',col_types = list(ch,ch,f,i,d))
# skip -> skiprows, n_max -> read maximum n_max rows
summary(gpa)
str(gpa)
```

## Data.table

 - key: speed
 - Data manipulation in R
 - **fread()**
 
```{r}
library(data.table)

#fread automatically deals with missing colnames
fread("./gpa.csv")
# infers separators, types, and crazy fast
fread('gpa.csv',select=c("numgrade","credit"))
fread('gpa.csv',drop = c("grade","code"))
fread('gpa.csv',drop=1:2)
```

## Importing Excel data

```{r}
library(readxl)
excel_sheets("gpa.xlsx")
g2020s = read_excel("gpa.xlsx",sheet=1)
g2019f = read_excel("gpa.xlsx",sheet=2)
g2019f


#using lapply
gpa2y = lapply(excel_sheets("gpa.xlsx"),read_excel, path = "gpa.xlsx")
gpa2y
str(gpa2y)
```
 
 
 