---
title: "Intermediate R"
author: "Sungbae Chun"
date: "5/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Relational Operators


```{r rel, echo=T}
# comparisons ==,!=,>,<,>=,<=
# are the same as python
l = c(10,3,5,6,8)
l >= 5
l[l>5]
v = c(3,5,6,1,9)
#elemenwise comparison
l>v
```

### Logical operators
`&, |, !`<br>
`T and T` won't work<br>
`vec and vec` - elementwise
```{r log}
T & T

```

### Conditional statements

```{r if}
x=1
if (x<1){
  print("0 is less than 1")
}else if (x==1){
  print("x is zero")
} else{
  print(paste("x is larger than",1))
}

```

## Loops

### While loop

```{r,eval=F}
while(condition){
  expr
  if (condition){
    break
  }
}
```
  
### For loop
```{r, eval=F}
for(var in seq){
  expr
  if (cond){
    next
    #next == continue
  }
}
```

## Functions

```{r, eval=F}
?func

my_fun = function(a1, a2 = a20){
  body
  return(some_val)
}
```

## Packages

```{r,eval=F}
install.packages(<pkgname>)
install.packages("ggvis")
```


```{r}
library("ggvis")
search()
```


### `lapply`

```{r}
a = list("a",2,c(10,20))
lapply(a,class)

n = c("Sungbae","Max","Win")
lapply(n,nchar)

multiply = function(x,fac){
  x*fac
}
v = c(1,2,3)

lapply(v,multiply,fac=3)
```

### `sapply`

lapply - returns list<br>
sapply - simplest struct
```{r}
n = c("Sungbae","Max","Win")
sapply(n,nchar)
sapply(n,nchar,USE.NAMES = F)
```
There are cases when sapply == lapply, you can check it using `identical(l,s)`

### `vapply`

 - **explicitly specify** output format

```{r}
n = c("Sungbae","Max","Win")

sapply(n,nchar)
vapply(n,nchar, numeric(1))

f_l = function(name){
  letters = strsplit(name, split = "")[[1]]
  return(c(first=letters[1],second=letters[2]))
}
sapply(n,f_l)
vapply(n,f_l,character(2))

```

## Useful functions

```{r}
li <- list(
  log = T,
  ch = "hello",
  invec = sort(rep(seq(8,2,by=-2),times=2))
)

seq(8,2,by = -2)
seq(1,10,by=2)

rep(c(1,2), times=2)
sort(rep(c(1,2), times=2))

str(li)

is.list(li)

unlist(li)

rev(li)

append(li,c(10,20))
```

### Regex

`grep, grepl, sub, gsub`

`grepl(pattern = <regex>, x = <str>)`

```{r}
n = c("Chun","Mametkulov")
grepl(pattern="a", x = n)
# ^ - start
grepl(patter="^C",x=n)
# $ - end
grepl(pattern="n$",x=n)

grep(pattern="n$",x=n)
```

`sub(pattern=<reg>, replacement=<str>, x = <str>)`
sub - only the first instance<br>

gsub - all
```{r}
n = c("sungbae","mametkulov")
gsub(pattern="u|a", replacement="_",
     x=n)
```


## Times and Dates

```{r}
today = Sys.Date()
today

class(today)

now = Sys.time()
now
class(now)

my_date = as.Date("1999-10-10")
my_date

my_date = as.Date("10/10/1999",format = "%d/%m/%Y")
my_date
class(my_date)

my_time = as.POSIXct("1999-10-10 11:25:15")
my_time

my_date
my_date+30
unclass(my_date)
format(my_date, "%A")

my_time
my_time+2000
unclass(my_time)
```
